---
title: "Vignette Title"
author: "Vignette Author"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

There are currently three supported cases to generate question metadata:

1. You want to prepare question metadata without images, which is done like 
  this:
```{r, eval=FALSE}
generate_question_jsons(projects_root = "the directory in which the projects 
  live", project_name = "project_name", has_images = FALSE)
```
  
2. You want to prepare question metadata with images, which is done like 
  this:
```{r, eval=FALSE}
generate_question_jsons(projects_root = "the directory in which the projects 
  live", project_name = "project_name", has_images = TRUE)
```
3. You want to prepare question metadata with images that are not sorted yet, 
  which is done like this:
```{r, eval=FALSE}
generate_question_jsons(projects_root = "the directory in which the projects 
  live", project_name = "project_name", has_images = TRUE)
sort_images(pathInImages = "project_root/project_name/questions/images_unsorted", 
  pathOutImages = "project_root/project_name/questions/out",
  pathXlsxFile = "project_root/project_name/questions/excelfile.xlsx"))
```
Note that the excelfile should be named as $"project_name.xlsx"$.
